Reducing retrieval error in MAIAC AOD with the `Earth_obs_cleaning` machine-learning workflow: Results from CONUS

### 2023-08-26 writing/CONUS_AOD.qmd
Caption DV density figure, plot descriptives, and caption overall CV table.

```{r basics}
source.files = "globals libraries util data modeling paper_functions"
for (filename in strsplit(source.files, " ")[[1]])
     source(sprintf("code/%s.R", filename))
suppressPackageStartupMessages(
   {library(patchwork)
    library(ggridges)
    library(flextable)
  })
the.sat = tools::toTitleCase(Wf$satellite)
IG = knitr::include_graphics
```

@@

<!--# any missingness in predictors to report?-->

After collocation there were `r scales::comma(nrow(dt))` matched AOD retrievals from `r uniqueN(dt$site)` AERONET sites in CONUS. The quartiles of the absolute difference from the labeled overpass time (one per 1200-km wide tile) and the time of the matched AERONET observation were `r tqs = round(quantile(dt[, abs(as.numeric(difftime(units = "secs", time.sat, time.ground)))], c(.25, .5, .75))); tqs["25%"]` s, `r tqs["50%"]` s, and `r tqs["75%"]` s. While the distribution of AOD is skewed, our estimate of retrieval error and the dependent variable for our predictive model, *diff~AOD~*, was fairly symmetric and Gaussian.

```{r describe_dv}
summary(dt[, list(y.diff)])
```

```{r plot_dv}
ggplot(dt, aes(x = y.diff)) +
  geom_histogram(binwidth = 0.005) +
  scale_x_continuous("MAIAC AOD minus AERONET AOD") +
  coord_cartesian(xlim = c(-1, 1)) +
  labs(title = paste0("Histogram of difference in collocated AOD (n=", scales::comma(dt[, .N]), ")"),
      caption = paste0("Not shown: ", dt[y.diff < -1 | y.diff > 1, .N], " AOD differences outside [-1, 1]")) +
  theme_bw()
```

Table \## includes the performance metrics for the full cross-validated training data set. Here **raw** is MCD19A2.V006 MAIAC AOD 470nm; **ground** is AERONET AOD interpolated to 470nm; and **corrected** is MAIAC after applying an XGBoost-modeled correction term.

```{r}
flextable(pretty.table.numbers(tar_read(cv.summary)[1, -"Year"]))
```

>>>>

```{r basics}
source.files = "globals libraries util data modeling paper_functions"
for (filename in strsplit(source.files, " ")[[1]])
     source(sprintf("code/%s.R", filename))
suppressPackageStartupMessages(
   {library(patchwork)
    library(gridExtra)
    library(flextable)
  })
the.sat = tools::toTitleCase(Wf$satellite)
IG = knitr::include_graphics
```

@@

After collocation there were `r scales::comma(nrow(dt))` matched AOD retrievals from `r uniqueN(dt$site)` AERONET sites in CONUS. The quartiles of the absolute difference from the labeled overpass time (one per 1200-km wide tile) and the time of the matched AERONET observation were `r tqs = round(quantile(dt[, abs(as.numeric(difftime(units = "secs", time.sat, time.ground)))], c(.25, .5, .75))); tqs["25%"]` s, `r tqs["50%"]` s, and `r tqs["75%"]` s. While the distribution of AOD is skewed, our estimate of retrieval error and the dependent variable for our predictive model, *diff~AOD~*, was fairly symmetric and Gaussian (see @fig-dv-density).

```{r plot_dv}
#| label: fig-dv-density
#| fig-cap: !expr caption
caption = with(list(d = dt), sprintf(
    "Histogram of the difference in collocated AOD (n=%s). Not shown: %s %s",
    scales::comma(d[, .N]),
    d[y.diff < -1 | y.diff > 1, .N],
    "AOD differences outside [-1, 1]"))
p1 <- ggplot(dt, aes(x = y.diff)) +
  geom_histogram(binwidth = 0.005) +
  coord_cartesian(xlim = c(-1, 1), expand = FALSE) +
  xlab("MAIAC AOD minus AERONET AOD") +
  theme_classic()
t1 <- summary(dt[, list(y.diff)], digits = 2)
colnames(t1) <- ""
t1 <- tableGrob(t1, theme=ttheme_minimal(), rows=NULL)  # transform into a tableGrob
p1 + inset_element(t1, left = 0.02, bottom = 0.6, right = 0.25, top = .98)
```

@tbl-cv-overall includes the performance metrics for the full cross-validated training data set. Here **raw** is MCD19A2.V061 MAIAC AOD 470nm; **ground** is AERONET AOD interpolated to 470nm; and **corrected** is MAIAC after applying an XGBoost-modeled correction term.

```{r}
#| label: tbl-cv-overall
#| tbl-cap: Overall performance from cross-validation.
flextable(pretty.table.numbers(tar_read(cv.summary)[1, -"Year"]))
```
